### 2020-06-16 report/sections/background/sessiontypes.tex
Report: minor background fix

\item \textbf{Polyadic communication}: 
More than one values can be communicated at once.
We refer to these as a \textit{vector} of values.

@@

\begin{prooftree}
\AxiomC{$\exists	j \in I. l_j = l$}
\AxiomC{$e_1 \downarrow v_1 \dots e_n \downarrow v_n$}
\AxiomC{$\mrole{p} \neq \mrole{q}$}
\RightLabel{\rulename{Comm}}
\TrinaryInfC{$
\mrole{p} :: \sel{q}{l}{e_1 \dots e_n}.~P \mid 
\mrole{q} :: \bra{p}{l_i(x_1 \dots x_n): Q_i}{i\in I} 
~\longrightarrow~ 
\ptprocess{p}{P} \mid \ptprocess{q}{Q_j[v_k/x_k]^n_{k=1}}
$}
\end{prooftree}

@@

P,Q ::= & & \text{Processes} \\
     & \mathbf{0} & \text{Nil Process} \\
\mid & \sel{p}{l}{e_1 \dots e_n}.~P & \text{Selection} \\
\mid & \bra{p}{l_i(x_1 \dots x_n): P_i}{i\in I} & \text{Branching} \\
\mid & \pcond{e}{P}{Q} & \text{Conditional} \\
\mid & \precursion{P} & \text{Recursive Process} \\
\mid & \precvar & \text{Process Variable} \\

@@

{Session types} represent the type theory for our session calculus.
We define the syntax of session types for binary sessions in
\cref{fig:bst}.

@@

T ::= & & \text{Session Types} \\
     & \mathbf{end} & \text{Termination} \\
\mid & \tbra{\role{p}}{l_i(U_1 \dots U_n):T_i}{i\in I} & \text{Branching} \\
\mid & \tsel{\role{p}}{l_i(U_1 \dots U_n):T_i}{i\in I} & \text{Selection} \\
\mid & \trec{T} & \text{Recursive Type} \\
\mid & \trecvar & \text{Type Variable} \\
\end{array}
\]

@@

\begin{prooftree}
\AxiomC{$
\Gamma \vdash e_1 : U_1 \dots \Gamma \vdash e_n : U_n
$}
\AxiomC{$\Gamma \vdash P : T$}
\RightLabel{\rulename{Ty-Sel}}
\BinaryInfC{$
\Gamma \vdash \sel{p}{l}{e_1 \dots e_n}.~P : 
\tselone{p}{l(U_1 \dots U_n)}: T
$}
\end{prooftree}

\begin{prooftree}
\AxiomC{$
\forall i \in I. ~ \left(
\Gamma, x_1: U_1, \dots, x_n: U_n \vdash P_i : T_i
\right)
$}
\RightLabel{\rulename{Ty-Bra}}
\UnaryInfC{$
\Gamma \vdash \bra{p}{l_i(x_1 \dots x_n): P_i}{i\in I} : 
\tbra{p}{l_i(U_1 \dots U_n):S_i}{i\in I}
$}
\end{prooftree}

@@

% Row height in tables
\renewcommand{\arraystretch}{1.6}
\begin{center}
\begin{tabular}{rcl}
$\dual{\tend}$ & = & $\tend$ \\
$\dual{\tbra{p}{l_i: T_i}{i\in I}}$ & = & 
	$\tsel{q}{l_i: \dual{T_i}}{i\in I}$ \\
$\dual{\tsel{p}{l_i: T_i}{i\in I}}$ & = & 
	$\tbra{q}{l_i: \dual{T_i}}{i\in I}$ \\
$\dual{\trec{T}}$ & = & $\trec{\dual{T}}$ \\
$\dual{\trecvar}$ & = & $\trecvar$ \\ 
\end{tabular}
\captionof{table}{Duality of Binary Session Types involving participants $\role{p}$ and $\role{q}$}
\label{table:dual}
\end{center}
\renewcommand{\arraystretch}{1}

@@

\[
\begin{array}{rl}
T_{\text{Client}} &= 
	\tselone{Server}{\tmsg{Succ(nat)}}.~
	\tbraone{Server}{\tmsg{Succ(int)}}.~
	\tend \\
T_{\text{Server}} &= \tbraone{Client}{
\begin{cases}
	\tmsg{Succ(int)}: \tselone{Client}{\tmsg{Succ(int)}} \\
	\tmsg{Quit()}:  \tend \\
\end{cases}
} \\
\end{array}
\]

@@

This motivates the concept of \textit{subtypes} in session types,
which allows a process to be typed by its ``supertype'' when required. 
$\subtype$\footnote{
The $\subtype$ operator is also an 
overloaded relation on 
sorts to express subsorting, 
i.e. $\texttt{nat} \subtype \texttt{int}$.
} defines the subtyping relation: 
$T \subtype T'$ reads \textit{$T$ is a subtype of $T'$},
and is defined coinductively on the structure of session types.

@@

\begin{itemize}
\item \textbf{Branching}: 
The supertype of a branching process offers 
a subset of the branches and 
expects more specific types of payload; 
intuitively, if a process expects to 
receive an \texttt{int}, 
it can handle a \texttt{nat} payload.

\item \textbf{Selection}: 
The supertype of a selection process offers a 
superset of the internal choices and 
can send more generic types of payload; 
intuitively, if a process sends a \texttt{nat}, 
the payload is compatible with receivers expecting 
a more generic \texttt{int} payload.
\end{itemize}

\begin{prooftree}
\AxiomC{$\forall i \in I.$}
\AxiomC{$U'_1 \subtype U_1 ~ \dots ~ U'_n \subtype U_n$}
\AxiomC{$S_i \subtype T'_i$}
\RightLabel{\rulename{Sub-Bra}}
\doubleLine
\TrinaryInfC{$
	\tbra{p}{l_i(U_1 \dots U_n):T_i}{i\in I \cup J}
	\subtype
	\tbra{p}{l_i(U'_1 \dots U'_n):T'_i}{i\in I}
$}
\end{prooftree}

\begin{prooftree}
\AxiomC{$\forall i \in I.$}
\AxiomC{$U_1 \subtype U'_1 ~ \dots ~ U_n \subtype U'_n$}
\AxiomC{$T_i \subtype T'_i$}
\RightLabel{\rulename{Sub-Sel}}
\doubleLine
\TrinaryInfC{$
	\tsel{p}{l_i(U_1 \dots U_n):T_i}{i\in I}
	\subtype
	\tsel{p}{l_i(U'_1 \dots U'_n):T'_i}{i\in I \cup J}
$}
\end{prooftree}

>>>>

\item \textbf{Polyadic communication}: 
More than one value can be communicated at once.
We refer to these as a \textit{vector} of values, $\vec v$.

@@

\begin{prooftree}
\AxiomC{$\exists	j \in I. l_j = l$}
\AxiomC{$\vec x_j = x_1, \dots, x_n$}
\AxiomC{$e_1 \downarrow v_1 \dots e_n \downarrow v_n$}
\AxiomC{$\mrole{p} \neq \mrole{q}$}
\RightLabel{\rulename{Comm}}
\QuaternaryInfC{$
\mrole{p} :: \sel{q}{l}{e_1 \dots e_n}.~P \mid 
\mrole{q} :: \bra{p}{l_i(\vec x_i): Q_i}{i\in I} 
~\longrightarrow~ 
\ptprocess{p}{P} \mid \ptprocess{q}{Q_j[v_k/x_k]^n_{k=1}}
$}
\end{prooftree}

@@

P,Q ::= & & \text{Processes} \\
     & \mathbf{0} & \text{Nil Process} \\
\mid & \sel{p}{l}{\vec e}.~P & \text{Selection} \\
\mid & \bra{p}{l_i(\vec x_i): P_i}{i\in I} & \text{Branching} \\
\mid & \pcond{e}{P}{Q} & \text{Conditional} \\
\mid & \precursion{P} & \text{Recursive Process} \\
\mid & \precvar & \text{Process Variable} \\

@@

{Session types} represent the type theory for our session calculus.
We define the syntax of session types for binary sessions in
\cref{fig:bst}. $\vec U$ denotes a vector of sorts.

@@

T ::= & & \text{Session Types} \\
     & \tend & \text{Termination} \\
\mid & \tbra{p}{l_i(\vec U_i):T_i}{i\in I} & \text{Branching} \\
\mid & \tsel{p}{l_i(\vec U_i):T_i}{i\in I} & \text{Selection} \\
\mid & \trec{T} & \text{Recursive Type} \\
\mid & \trecvar & \text{Type Variable} \\
\end{array}
\]

@@

\begin{prooftree}
\AxiomC{$
\Gamma \vdash e_1 : U_1 ~ \dots ~ \Gamma \vdash e_n : U_n
$}
\AxiomC{$\Gamma \vdash P : T$}
\RightLabel{\rulename{Ty-Sel}}
\BinaryInfC{$
\Gamma \vdash \sel{p}{l}{e_1 \dots e_n}.~P : 
\tselone{p}{l(U_1, \dots, U_n)}: T
$}
\end{prooftree}

\begin{prooftree}
\AxiomC{$
\forall i \in I \left(
\vec x_i = x_1, \dots, x_n
\qquad
\vec U_i = U_1, \dots, U_n
\qquad
\Gamma, x_1: U_1, \dots, x_n: U_n \vdash P_i : T_i
\right)$}
\RightLabel{\rulename{Ty-Bra}}
\UnaryInfC{$
\Gamma \vdash \bra{p}{l_i(\vec x_i): P_i}{i\in I} : 
\tbra{p}{l_i(\vec U_i):S_i}{i\in I}
$}
\end{prooftree}

@@

% Row height in tables
\renewcommand{\arraystretch}{1.6}
\begin{center}
\begin{tabular}{rcl}
$\dual{\tend}$ & = & $\tend$ \\
$\dual{\tbra{p}{l_i: T_i}{i\in I}}$ & = & 
	$\tsel{q}{l_i: \dual{T_i}}{i\in I}$ \\
$\dual{\tsel{p}{l_i: T_i}{i\in I}}$ & = & 
	$\tbra{q}{l_i: \dual{T_i}}{i\in I}$ \\
$\dual{\trec{T}}$ & = & $\trec{\dual{T}}$ \\
$\dual{\trecvar}$ & = & $\trecvar$ \\ 
\end{tabular}
\captionof{table}{Duality of Binary Session Types 
Involving Participants $\role{p}$ and $\role{q}$}
\label{table:dual}
\end{center}
\renewcommand{\arraystretch}{1}

@@

\[
\begin{array}{rl}
T_{\text{Client}} &= 
	\tselone{Server}{\tmsg{Succ(nat)}}.~
	\tbraone{Server}{\tmsg{Succ(int)}}.~
	\tend \\
T_{\text{Server}} &= \tbraone{Client}{
\begin{cases}
	\tmsg{Succ(int)}: \tselone{Client}{\tmsg{Succ(int)}}. \tend \\
	\tmsg{Quit()}:  \tend \\
\end{cases}
} \\
\end{array}
\]

@@

This motivates the concept of \textit{subtypes} in session types,
which allows a process to be typed by its ``supertype'' when required. 
$\subtype$\footnote{
The $\subtype$ operator is also an 
overloaded relation on 
sorts to express subsorting, 
i.e. $\texttt{nat} \subtype \texttt{int}$.
} defines the subtyping relation: 
$T \subtype T'$ reads \textit{$T$ is a subtype of $T'$},
and is defined coinductively on the structure of $T$.

@@

\begin{itemize}
\item \textbf{Selection}: 
The supertype of a selection process offers a 
superset of the internal choices and 
can send more generic types of payload; 
intuitively, if a process sends a \texttt{nat}, 
the payload is compatible with receivers expecting 
a more generic \texttt{int} payload.

\item \textbf{Branching}: 
The supertype of a branching process offers 
a subset of the branches and 
expects more specific types of payload; 
intuitively, if a process expects to 
receive an \texttt{int}, 
it can handle a \texttt{nat} payload.
\end{itemize}

\begin{prooftree}
\AxiomC{$\forall i \in I.\left(
\vec U_i = U_1, \dots, U_n 
\quad
\vec U'_i = U'_1, \dots, U'_n
\quad
U_1 \subtype U'_1 ~ \dots ~ U_n \subtype U'_n
\quad
T_i \subtype T'_i
\right)$}
\RightLabel{\rulename{Sub-Sel}}
\doubleLine
\UnaryInfC{$
	\tsel{p}{l_i(\vec U_i):T_i}{i\in I}
	\subtype
	\tsel{p}{l_i(\vec U'_i):T'_i}{i\in I \cup J}
$}
\end{prooftree}

\begin{prooftree}
\AxiomC{$\forall i \in I.\left(
\vec U_i = U_1, \dots, U_n 
\quad
\vec U'_i = U'_1, \dots, U'_n
\quad
U'_1 \subtype U_1 ~ \dots ~ U'_n \subtype U_n
\quad
T_i \subtype T'_i
\right)$}
\RightLabel{\rulename{Sub-Bra}}
\doubleLine
\UnaryInfC{$
	\tbra{p}{l_i(\vec U_i):T_i}{i\in I \cup J}
	\subtype
	\tbra{p}{l_i(\vec U'_i):T'_i}{i\in I}
$}
\end{prooftree}
